---
title: "Homework 9"
author: "Gabriel Lahman"
date: "4/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("finite_markov_chain_functions.R")
```

## Question 1

#### 1.1

```{r}
data("preproglucacon")
bases = preproglucacon[,2]
est_trans_mat = estimate_transition_matrix(bases)
```
```{r echo=FALSE}
print(est_trans_mat)
```

#### 1.2

```{r}
mot = mean_occupancy_time_fun(bases)
```
```{r echo = FALSE}
print(mot)
```


#### 1.3

```{r}

n = 100
x = matrix(nrow = n, ncol = 4)
x[1,] = c(1,0,0,0)
d = rep(NA, n)
for (i in 2:n) {
  x[i,] = x[i - 1,] %*% est_trans_mat
  d[i] = sum(abs(x[i,] - x[i - 1,]))
}
plot(d, xlab = "n", main = "Dn")
```

The plot converges to zero incredibly fast, showing that we reach a stationary matrix in under 10 iterations.


#### 1.4

```{r}
v = mean_recurrence_time_fun(bases)
p = 1/v
```
```{r echo = FALSE}
print("p")
print(p)
print("x100")
print(x[100,])
```

The vectors are pratically equal.
## Question 2

Recurrent States: 1,5,6,4

Transient:3

Absorbing: 6

No ergodic components

Cyclic components: (6), (2,4), (1,5)



## Question 3


#### 3.1

$P(Y = (y_1, ..., y_8)) = P(Y_1 = y_1)P(Y_4 = y_4)P(Y_3 = y_3 | Y_1, Y_4)P(Y_5=y_5|Y_4)P(Y_2=y_2|Y_3)P(Y_7=y_7|Y_3, Y_5)P(Y_8=y_8|Y_5,Y_6,Y_7)$

#### 3.2

#### 3.3

## Question 4

#### 4.1

```{r}
f = function() {
  X = rexp(1)
  Y = rnorm(1)
  Z = rnorm(1, mean = Y, sd = 1/sqrt(X))
  return (c(X,Y,Z))
}
```

#### 4.2

$X|Y,Z ~ Gamma (\alpha = 3/2, \beta = \frac{z-y}{2})$

$Y|X,Z ~ Normal(\mu = zx, \sigma^2 = \frac{1}{1-x})$


#### 4.3

```{r}
f2 = function(state) {
  x1_y0z0 = rgamma(1, shape = 3/2, rate = (state[[2]] - state[[3]])/ 2)
  y1_x1z0 = rnorm(1, mean = x1_y0z0 * state[[3]], sd = sqrt(abs(1/1 - x1_y0z0)))
  z1_x1y1 = rnorm(1, mean = y1_x1z0, sd = 1 / sqrt(abs(x1_y0z0)))
  return (c(x1_y0z0, y1_x1z0, z1_x1y1))
}
```

#### 4.4

I am not sure my last function is correct as it is producing NA's, so I did not populate the data.
```{r}
state = f()

#for (i in 1:10000) {
#  print(state)
#  state = f2(state)
#}
```


## Question 5

#### 5.1

$E[Y_1] = E[\rho Y_0 + \epsilon_1] = \rho E[Y_0] + E[\epsilon_1] = \rho E[\epsilon_0] + 0 = \rho (0) = 0$

#### 5.2

$Var[Y_1] = Var[\rho Y_0 + \epsilon_1] = \rho^2 Var[Y_0] + \sigma^{2}_{\epsilon_1} = \rho^2 (1) + 1 = \rho^2 + 1$

#### 5.3

$E[Y_2] = E[\rho Y_1 + \epsilon_2] = \rho E[Y_1] + E[\epsilon_2] = \rho (0) + 0 = 0$

#### 5.4

$Var[Y_2] = Var[\rho Y_1 + \epsilon_2] = \rho^2 Var[Y_1] + \sigma^{2}_{\epsilon_2} = \rho^2 (\rho^2 + 1) + 1 = \rho^4 + \rho^2 + 1$

#### 5.5

$E[Y_3] = E[\rho Y_2 + \epsilon_3] = \rho E[Y_2] + E[\epsilon_3] = \rho (0) + 0 = 0$

#### 5.6

$Var[Y_3] = Var[\rho Y_2 + \epsilon_3] = \rho^2 Var[Y_2] + \sigma^{2}_{\epsilon_3} = \rho^2 (\rho^4 + \rho^2 + 1) + 1 = \rho^6 + \rho^4 + \rho^2 + 1$

## Question 6

#### 6.1

$E[Y_1] = E[\rho Y_0 + \phi \epsilon_0 + \epsilon_1] = \rho E[\epsilon_0] + \phi E[\epsilon_0] + E[\epsilon_1] = \rho (0) + \phi (0) = 0 = 0$

#### 6.2

$Var[Y_1] = Var[\rho Y_0 + \phi \epsilon_0 + \epsilon_1] = \rho^2 Var[Y_0] + \phi_2 Var[\epsilon_0] + Var[\epsilon_1] = \rho^2 + \phi^2 + 1$

#### 6.3

$E[Y_2] = E[\rho Y_1 + \phi \epsilon_1 + \epsilon_2] = \rho E[\epsilon_1] + \phi E[\epsilon_1] + E[\epsilon_2] = \rho (0) + \phi (0) = 0 = 0$

#### 6.4

$Var[Y_2] = Var[\rho Y_1 + \phi \epsilon_1 + \epsilon_2] = \rho^2 Var[Y_1] + \phi_2 Var[\epsilon_1] + Var[\epsilon_2] = \rho^2 (\rho^2 + \phi^2 + 1) + \phi^2 + 1 \\ = \rho^4 + \rho^2 \phi^2 + \rho^2 + \phi^2 + 1 = \rho^2(\rho^2 + 1) + \phi^2(\rho^2 + 1) + 1 = (\rho^2 + \phi^2)(\rho^2 + 1) + 1$

#### 6.5

$E[Y_3] = E[\rho Y_2 + \phi \epsilon_2 + \epsilon_3] = \rho E[\epsilon_2] + \phi E[\epsilon_2] + E[\epsilon_3] = \rho (0) + \phi (0) + 0 = 0$

#### 6.6

$Var[Y_3] = Var[\rho Y_2 + \phi \epsilon_2 + \epsilon_3] = \rho^2 Var[Y_2] + \phi^2 Var[\epsilon_2] + Var[\epsilon_3] \\ = \rho^2 ((\rho^2 + \phi^2)(\rho^2 + 1) + 1) + \phi^2 + 1 = \rho^2 (\rho^4 + \rho^2 + \rho^2 \phi^2 + \phi^2 + 1) + \phi^2 + 1 \\ = \rho^6 + \rho^4 + \rho^4 \phi^2 + \rho^2 \phi^2 + \rho^2 + \phi^2 + 1 = \rho^6 + \rho^4 + \rho^2 + \phi^2 (\rho^4 + \rho^2 + 1) + 1 \\ = \rho^2 (\rho^4 + \rho^2 + \rho^1) + \phi^2 (\rho^4 + \rho^2 + 1) + 1 \\ = (\rho^2 + \phi^2)(\rho^4 + \rho^2 + 1) + 1$